---
title: "Data 550 Final Project"
author: "Priya Edupuganti"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: yes
    toc_float:
      collapsed: true
---
```{r, here-i-am, include=FALSE}
here::i_am(
  "Edupuganti_550_final.Rmd"
)
```

```{r Chunk options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

## Introduction

### Goal
The goal of the current report is to illustrate descriptive statistic values associated with CTG Measurements, identify critical risk factors through linear regression modeling, and cross validate the results.

### Data
The dataset in use focuses on fetal health, containing cardiotocogram (CTG) measurements and assessing the health of 2126 fetuses. The data identifies a total of 21 different measurement variables that CTGs can provide. There's an additional ordinal variable, identifying fetuses' health status as "Normal", "Suspect", and "Pathological."

If a clear association between fetus health and a few of the measurements can be identified and trained to a model, CTGs' effectiveness in reducing child mortality rates can increase, and access to high-level practioners is not required to determine fetal health.

Source: Ayres de Campos et al. (2000) SisPorto 2.0 A Program for Automated Analysis of Cardiotocograms. J Matern Fetal Med 5:311-318. https://www.kaggle.com/datasets/andrewmvd/fetal-health-classification.

```{r, Loading Outputs}
Table <- readRDS(
  here::here("output/table.rds")
)

Plot <- readRDS(
  here::here("output/scatterplot.rds")
)
```

## Descriptive Statistics

### Summary Table of Critical Measurements
```{r, Table 1}
Table
```

The table above displays average Heart Rate, Fetal Movement, Uterine Contractions, and Abnormal Short-Term CTG Variability by fetal health status. Standard Deviations are located in parantheses next to each respective average, and the number of observations per status is noted as well.

### Relationship between Heart Rate and Movement
```{r, HR_Movement_Scatter}
Plot
```

The figure above displays fetal heart rate against fetal movements per second with points colored by health status. Clusters for Pathological and/or Suspect measurements can suggest possible relationships between fetal activity and health status.